### 2019-07-09 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Corrections fautes orthographes

Both D-UCB and SW-UCB build on a stationary policy, but for example \ExpThreeS{} from \cite{Auer02NonStochastic}
builds on the \ExpThree{} policy, which is designed for the adversarial case.
\ExpThreeS{} actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
It constitues a good baseline for the numerical experiments in Section~\ref{sec:6:NumericalExperiments},
even if we did find that \ExpThreeS{} performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
% , so we do not include it in .
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proven to be less efficient empirically, or are designed for more specific settings.

@@

\paragraph{Actively adaptive algorithms.}
%
% We distinguish two families,
A first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} used a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
%
The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but their approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
%
The \ExpThreeR{} algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with \ExpThree, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
and a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
The \CUSUM{} test is rather complicated and it is parametric: it uses the first $M$ samples (for a fixed $M\in\N^*$) from one arm to compute an average $\widehat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when one of the random walks crosses a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
and we found in numerical experiments that \PHT-\UCB{} performs very similarly to \CUSUMUCB.
Even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w\in 2\N^*$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametrized by $\omega\in(0,1)$ (note that it is called $\alpha$ in both papers, but we rename it $\omega$ in this chapter to avoid clutter with $\alpha$ of $\UCB_1$ in \eqref{eq:2:UCB_index}).
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\omega \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

\paragraph{Positionning of our results.}
%
The two algorithms \CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the break-points are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of break-points  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of the proposed algorithm.
% This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\Delta^{\text{change}} = \max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\Delta^{\text{opt}} = \max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.

@@

Before giving larger results that compare our approach against other algorithms, we consider the two problems $1$ and $2$ presented above in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} (for $T=5000$).
\emph{For one (random) simulation}, we give below examples of the efficiency of the change-point detection algorithm used by \GLRklUCB{} as well as \CUSUMklUCB{} and \MklUCB,
by showing the times at which they detect a change and reinitialize their memory of observations of one or all the arms.
% We also display the number of observations (since the last restart) of the arm on which the change was detected.
As shown below, in the experiments summary, the three algorithms based on change-point detection obtain sublinear regret, but they have different behaviors.
The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
%
On the examples presented below as well as large scale examples, all the tests appear to have no (or a small number of) false alarm, and always a small detection delay for changes that are ``easy enough'' to be efficiently detected.
% The four algorithms are presented in the order of decreasing performance (\ie, the first one is empirically the most efficient, that is, it obtains a smaller regret in average).

@@

Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} displays the locations of the successively detected change-points by four algorithms,
$1)$ \MklUCB,
$2)$ \CUSUMklUCB,
$3)$ \GLRklUCB{} with \emph{local restart},
$4)$ \GLRklUCB{} with \emph{global restart}.
In this ``easy'' problem, all algorithms correctly detect the important changes, except \CUSUM{} which detected a change on arm $1$ (\textcolor{red}{in red}) twice after the change-point located at $t=3000$.
The three other algorithms have a very small delay, for instance only $9$ samples from arm $0$ (\textcolor{blue}{in blue}) after its change at time $t=2000$ are enough for the \GLR{} test to detect a change.
Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, simply because the tuning of its forced exploration probability (\ie, $\omega$) makes it explore suboptimal arms more often.

@@

Finally, in the case of hard problems, like problems $3$ and $5$, that have a lot of changes but where the optimal arm barely changes, we verify in Figure~\ref{fig:6:meanRegretPb5} that \klUCB{} and TS can outperform the oracle policy.
Indeed the oracle policy is suboptimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).
We can also verify that the two options, \textbf{Local} and \textbf{Global} restart, for \GLRklUCB, give close results, and that the \textbf{Local} option is always better.

>>>>

Both D-UCB and SW-UCB build on a stationary policy, but for example \ExpThreeS{} from \cite{Auer02NonStochastic}
builds on the \ExpThree{} policy, which is designed for the adversarial case.
\ExpThreeS{} actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
It constitutes a good baseline for the numerical experiments in Section~\ref{sec:6:NumericalExperiments},
even if we did find that \ExpThreeS{} performs worse that the most of the other approaches based on extending stationary bandit algorithms (\eg, SW-UCB).
% , so we do not include it in .
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proven to be less efficient empirically, or are designed for more specific settings.

@@

\paragraph{Actively adaptive algorithms.}
%
% We distinguish two families,
A first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} used a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
%
The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but their approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
%
The \ExpThreeR{} algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with \ExpThree, and the history of all arms are reset as soon as a sub-optimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
and a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
The \CUSUM{} test is rather complicated and it is parametric: it uses the first $M$ samples (for a fixed $M\in\N^*$) from one arm to compute an average $\widehat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when one of the random walks crosses a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
and we found in numerical experiments that \PHT-\UCB{} performs very similarly to \CUSUMUCB.
Even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w\in 2\N^*$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parameterized by $\omega\in(0,1)$ (note that it is called $\alpha$ in both papers, but we rename it $\omega$ in this chapter to avoid clutter with $\alpha$ of $\UCB_1$ in \eqref{eq:2:UCB_index}).
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\omega \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

@@

\paragraph{Positioning of our results.}
%
The two algorithms \CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_k(t))$-- mostly saying that the break-points are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of break-points  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of the proposed algorithm.
% This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\Delta^{\text{change}} = \max_{k,t} |\mu_k(t) - \mu_k(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\Delta^{\text{opt}} = \max_{k,k'} |\mu_k(t) - \mu_{k'}(t)|$.

@@

Before giving larger results that compare our approach against other algorithms, we consider the two problems $1$ and $2$ presented above in Figures~\ref{fig:6:Problem_1} and \ref{fig:6:Problem_2} (for $T=5000$).
\emph{For one (random) simulation}, we give below examples of the efficiency of the change-point detection algorithm used by \GLRklUCB{} as well as \CUSUMklUCB{} and \MklUCB,
by showing the times at which they detect a change and reinitialize their memory of observations of one or all the arms.
% We also display the number of observations (since the last restart) of the arm on which the change was detected.
As shown below, in the experiments summary, the three algorithms based on change-point detection obtain sub-linear regret, but they have different behaviors.
The test in \MUCB{} usually detects less changes, while the test in \CUSUMUCB{} usually detects more changes and even unnecessary changes.
%
On the examples presented below as well as large scale examples, all the tests appear to have no (or a small number of) false alarm, and always a small detection delay for changes that are ``easy enough'' to be efficiently detected.
% The four algorithms are presented in the order of decreasing performance (\ie, the first one is empirically the most efficient, that is, it obtains a smaller regret in average).

@@

Figure~\ref{fig:6:Visualizing_locations_of_change_points_for_different_algorithms__4algs_Pb1} displays the locations of the successively detected change-points by four algorithms,
$1)$ \MklUCB,
$2)$ \CUSUMklUCB,
$3)$ \GLRklUCB{} with \emph{local restart},
$4)$ \GLRklUCB{} with \emph{global restart}.
In this ``easy'' problem, all algorithms correctly detect the important changes, except \CUSUM{} which detected a change on arm $1$ (\textcolor{red}{in red}) twice after the change-point located at $t=3000$.
The three other algorithms have a very small delay, for instance only $9$ samples from arm $0$ (\textcolor{blue}{in blue}) after its change at time $t=2000$ are enough for the \GLR{} test to detect a change.
Very small delays are possible only after having collected a lot of samples of the arm which changed, and for instance the same algorithm obtains a delay of $32$ samples from \textcolor{red}{arm $1$} for the next change at $t=3000$, because this arm was sampled less.
On this problem, while \MklUCB{} detects the same changes as \GLRklUCB, with larger but comparable delays, it is seen to obtain a larger regret than the two \GLR{} variants, simply because the tuning of its forced exploration probability (\ie, $\omega$) makes it explore sub-optimal arms more often.

@@

Finally, in the case of hard problems, like problems $3$ and $5$, that have a lot of changes but where the optimal arm barely changes, we verify in Figure~\ref{fig:6:meanRegretPb5} that \klUCB{} and TS can outperform the oracle policy.
Indeed the oracle policy is sub-optimal as it restarts as soon as one arm change but is unaware of the meaningful changes, and stationary policies which quickly identify the best arm will play it most of the times, achieving a smaller regret.
We note that, sadly, all actively adaptive policies fail to outperform stationary policies on such hard problems, because they do not observe enough rewards from each arm between two restarts (\ie, the Assumptions~\ref{ass:6:LongPeriodsGlobal} and \ref{ass:6:LongPeriods} for the two Theorems~\ref{thm:6:mainRegretBoundGlobal} and \ref{thm:6:mainRegretBound} are not satisfied).
We can also verify that the two options, \textbf{Local} and \textbf{Global} restart, for \GLRklUCB, give close results, and that the \textbf{Local} option is always better.

### 2019-07-09 3-Appendices/2-Appendix/appendix2.tex
Corrections fautes orthographes

\paragraph{Motivations for anytime algorithms.}
%
As introduced in Chapter~\ref{chapter:2},
an online reinforcement learning algorithm is \emph{anytime} if it does not need to know in advance the horizon $T$ of the experiment.
%
Depending on the context of the practical application of interest, it might be unrealistic to assume to know in advance $T$. We give two examples to illustrate where this prior knowledge can be realistic or not.
On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for this model we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an uplink and then downlink message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages. While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies \cite{Centenaro16}), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.

@@

\paragraph{Solution to ``patch'' a non-anytime algorithm.}
%
A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06} and used for instance in \cite{Auer10,AuerGajaneOrtner18}.
%
In the context of adversarial or stochastic multi-armed bandits,
the performance of an algorithm is measured by its regret,
and we study two families of sequences of growing horizons (geometric and exponential)
to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log(T))$.
We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log(T))$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
However, our study was only focussing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.

@@

To conclude this work, we would need to complete the study of the doubling tricks, and instead of focussing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bounds.
Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.
We are interested in pursuing this work, in the hope of finding an intermediate sequence, growing faster than a geometric but slower than an exponential doubling sequence, that can preserve both worlds, problem-dependent bounds in $\cO(\log(T))$ and problem-independent bounds in $\cO(\sqrt{T})$.

>>>>

\paragraph{Motivations for anytime algorithms.}
%
As introduced in Chapter~\ref{chapter:2},
an online reinforcement learning algorithm is \emph{anytime} if it does not need to know in advance the horizon $T$ of the experiment.
%
Depending on the context of the practical application of interest, it might be unrealistic to assume to know in advance $T$. We give two examples to illustrate where this prior knowledge can be realistic or not.
On the one hand, consider clinical trials: it is likely that the number of patients is known before starting the trial, and for this model we refer to all the research that studies the setting of fixed-time best-arm identification \cite{audibert2010best,Garivier16BAI}.
On the second hand, consider cognitive radio and decentralized MAB learning implemented on IoT devices (like we present it in Chapter~\ref{chapter:4}): usually a learning step corresponds to an up-link and then down-link message sent and received by the IoT device, and so the time horizon $T$ denotes the total number of such messages. While it can be assumed that the device will run for instance for $10$ years (as it is advertised by some companies \cite{Centenaro16}), many kinds of application such as medical sensors or connected fields cannot predict the number of total communications before setting up the device.

@@

\paragraph{Solution to ``patch'' a non-anytime algorithm.}
%
A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06} and used for instance in \cite{Auer10,AuerGajaneOrtner18}.
%
In the context of adversarial or stochastic multi-armed bandits,
the performance of an algorithm is measured by its regret,
and we study two families of sequences of growing horizons (geometric and exponential)
to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log(T))$.
We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log(T))$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
However, our study was only focusing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.

@@

To conclude this work, we would need to complete the study of the doubling tricks, and instead of focusing on two families (geometric and exponential), we need to completely characterize the doubling tricks that allow to preserve certain regret bounds.
Such doubling is given either by the function mapping the current time $t$ to the current estimate of the horizon $T(t)$, or the sequence $(T_i)_{i\in\N^*}$ of successive estimates of the horizons.
We are interested in pursuing this work, in the hope of finding an intermediate sequence, growing faster than a geometric but slower than an exponential doubling sequence, that can preserve both worlds, problem-dependent bounds in $\cO(\log(T))$ and problem-independent bounds in $\cO(\sqrt{T})$.
